---
title: "Chapter6-7"
author: "Li Sun"
date: "September 26, 2017"
output: html_document
---

## Chapter6 Scales, axes and legends  
### 6.1 Introduction  

### 6.2 Modifying scales  
```{r}
library(ggplot2)
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

##Default
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_discrete()


##override some default  
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous("A really awesome x axis label") +
  scale_y_continuous("An amazingly great y axis label")


ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous("Label 1") +
  scale_x_continuous("Label 2")

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous("Label 2")

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_sqrt() +
  scale_colour_brewer()
```  
scale_The name of the aesthetic (e.g., colour, shape or x)  

or  

scale_The name of the scale (e.g., continuous, discrete, brewer).  

### 6.3 Guides: legends and axes  
Guides allow you to
read observations from the plot and map them back to their original values.  

#### 6.3.1 Scale title  
```{r}
df <- data.frame(x = 1:2, y = 1, z = "a")
p <- ggplot(df, aes(x, y)) + geom_point()
p + scale_x_continuous("X axis")
p + scale_x_continuous(quote(a + mathematical ^ expression))

p <- ggplot(df, aes(x, y)) + geom_point(aes(colour = z))
p +
  xlab("X axis") +
  ylab("Y axis")
p + labs(x = "X axis", y = "Y axis", colour =
           "Colour\nlegend")

## no label
p <- ggplot(df, aes(x, y)) + 
  geom_point() +
  theme(plot.background = element_rect(colour = "grey50"))
p + labs(x = "", y = "")
p + labs(x = NULL, y = NULL)
```  


#### 6.3.2 Breaks and labels  
The breaks argument controls which values appear as tick marks on axes and keys on legends.  
```{r}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)
axs <- ggplot(df, aes(x, y)) +
  geom_point() +
  labs(x = NULL, y = NULL)
axs
axs + scale_x_continuous(breaks = c(2000, 4000))
axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))  



leg <- ggplot(df, aes(y, x, fill = x)) +
  geom_tile() +
  labs(x = NULL, y = NULL)
leg
leg + scale_fill_continuous(breaks = c(2000, 4000))
leg + scale_fill_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```  

If you want to relabel the breaks in a categorical scale, you can use a named labels vector:  
```{r}
df2 <- data.frame(x = 1:3, y = c("a", "b", "c"))
ggplot(df2, aes(x, y)) +
  geom_point()
ggplot(df2, aes(x, y)) +
  geom_point() +
  scale_y_discrete(labels = c(a = "apple", b = "banana", c = "carrot"))

axs + scale_x_continuous(breaks = NULL)
axs + scale_x_continuous(labels = NULL)

leg + scale_fill_continuous(breaks = NULL)
leg + scale_fill_continuous(labels = NULL)
```  

The scales package provides a number of useful labelling functions:  
```{r}
axs + scale_y_continuous(labels = scales::percent_format())
axs + scale_y_continuous(labels = scales::dollar_format("$"))
leg + scale_fill_continuous(labels = scales::unit_format("k", 1e-3))
```  


```{r}
df <- data.frame(x = c(2, 3, 5, 10, 200, 3000), y = 1)
ggplot(df, aes(x, y)) +
  geom_point() +
  scale_x_log10()
mb <- as.numeric(1:10 %o% 10 ^ (0:4))
ggplot(df, aes(x, y)) +
  geom_point() +
  scale_x_log10(minor_breaks = log10(mb))
```  

### 6.4 Legends  
#### 6.4.1 Layers and legends  

```{r}
df <- data.frame(x=0,y = c(1, 2, 3), z=c('a','b','c'))
ggplot(df, aes(y, y)) +
  geom_point(size = 4, colour = "grey20") +
  geom_point(aes(colour = z), size = 2)
ggplot(df, aes(y, y)) +
  geom_point(size = 4, colour = "grey20", show.legend = TRUE) +
  geom_point(aes(colour = z), size = 2)

norm <- data.frame(x = rnorm(1000), y = rnorm(1000))
norm$z <- cut(norm$x, 3, labels = c("a", "b", "c"))
ggplot(norm, aes(x, y)) +
  geom_point(aes(colour = z), alpha = 0.1)
ggplot(norm, aes(x, y)) +
  geom_point(aes(colour = z), alpha = 0.1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

ggplot(df, aes(x, y)) + geom_point(aes(colour = z))
ggplot(df, aes(x, y)) + geom_point(aes(shape = z))
ggplot(df, aes(x, y)) + geom_point(aes(shape = z, colour = z))
```  


#### 6.4.2 Legend layout  
```{r}
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
base <- ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z), size = 3) +
  xlab(NULL) +
  ylab(NULL)
base + theme(legend.position = "right") # the default
base + theme(legend.position = "bottom")
base + theme(legend.position = "none")
```
`legend.direction`: layout of items in legends (“horizontal” or “vertical”)
`legend.box`: arrangement of multiple legends (“horizontal” or “vertical”)
`legend.box.just`: justification of each legend within the overall bounding box, when there are multiple legends (“top”, “bottom”, “left”, or “right”)  
```{r}
base <- ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z), size = 3)
base + theme(legend.position = c(0, 1), legend.justification = c(0, 1))
base + theme(legend.position = c(0.5, 0.5), legend.justification = c(0.5, 0.5))
base + theme(legend.position = c(1, 0), legend.justification = c(1, 0))
```  


#### 6.4.3 Guide functions  
```{r}
df <- data.frame(x = 1, y = 1:3, z = 1:3)
base <- ggplot(df, aes(x, y)) + geom_raster(aes(fill = z))
base
base + scale_fill_continuous(guide = guide_legend())
base + guides(fill = guide_legend())
```  

##### 6.4.3.1 `guide_legend()`  
The legend guide displays individual keys in a table.  

```{r}
df <- data.frame(x = 1, y = 1:4, z = letters[1:4])
# Base plot
p <- ggplot(df, aes(x, y)) + geom_raster(aes(fill = z))
p
p + guides(fill = guide_legend(ncol = 2))
p + guides(fill = guide_legend(ncol = 2, byrow = TRUE))

p <- ggplot(df, aes(1, y)) + geom_bar(stat = "identity", aes(fill = z))
p
p + guides(fill = guide_legend(reverse = TRUE))
```  


##### 6.4.3.2 `guide_colourbar`  
```{r}
df <- data.frame(x = 1, y = 1:4, z = 4:1)
p <- ggplot(df, aes(x, y)) + geom_tile(aes(fill = z))
p
p + guides(fill = guide_colorbar(reverse = TRUE))
p + guides(fill = guide_colorbar(barheight = unit(4, "cm")))
```  


### 6.5 Limits  

```{r}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + geom_point()
base
base + scale_x_continuous(limits = c(1.5, 2.5))
#> Warning: Removed 2 rows containing missing values (geom_point).
base + scale_x_continuous(limits = c(0, 4))
base + xlim(0, 4)
base + xlim(4, 0)
base + lims(x = c(0, 4))
```  

To eliminate this space, set expand = c(0, 0). This is useful in conjunction with geom_raster():  
```{r}
ggplot(faithfuld, aes(waiting, eruptions)) + 
  geom_raster(aes(fill = density)) +
  theme(legend.position = "none")
ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position = "none")

```  
 
 
The default is scales::censor() which replaces any value outside the limits with NA. Another option is scales::squish() which squishes all values into the range:
```{r}
df <- data.frame(x = 1:5)
p <- ggplot(df, aes(x, 1)) + geom_tile(aes(fill = x), colour = "white")
p
p + scale_fill_gradient(limits = c(2, 4))
p + scale_fill_gradient(limits = c(2, 4), oob = scales::squish)
```  


### 6.6 Scales toolbox  
#### 6.6.1 Continuous position scales  
```{r}
# Convert from fuel economy to fuel consumption
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(trans = "reciprocal")
# Log transform x and y axes
ggplot(diamonds, aes(price, carat)) +
  geom_bin2d() +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")


base <- ggplot(economics, aes(date, psavert)) +
  geom_line(na.rm = TRUE) +
  labs(x = NULL, y = NULL)
base # Default breaks and labels
base + scale_x_date(date_labels = "%y", date_breaks = "5 years")

base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2005-01-01")),
  date_labels = "%b %y",
  date_minor_breaks = "1 month"
  )
base + scale_x_date(
  limits = as.Date(c("2004-01-01", "2004-06-01")),
  date_labels = "%m/%d",
  date_minor_breaks = "2 weeks"
  )


```  


#### 6.6.2 Color  
We’ll use a modern attempt called the `hcl` colour
space, which has three components of h ue, c hroma and l
uminance:  

`Hue` is a number between 0 and 360 (an angle) which gives the “colour” of the colour: like blue, red, orange, etc.  

`Luminance` is the lightness of the colour. A luminance of 0 produces black, and a luminance of 1 produces white.  

`Chroma` is the purity of a colour. A chroma of 0 is grey, and the maximum value of chroma varies with luminance.  

##### 6.6.2.1 Continuous  
```{r}
erupt <- ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_raster() +
  scale_x_continuous(NULL, expand = c(0, 0)) +
  scale_y_continuous(NULL, expand = c(0, 0)) +
  theme(legend.position = "none")

erupt
erupt + scale_fill_gradient(low = "white", high = "black")
erupt + scale_fill_gradient(
  low = munsell::mnsl("5G 9/2"),
  high = munsell::mnsl("5G 6/8")
  )


mid <- median(faithfuld$density)
erupt + scale_fill_gradient2(midpoint = mid)

erupt + scale_fill_gradientn(colours = terrain.colors(7))
erupt + scale_fill_gradientn(colours = colorspace::heat_hcl(7))
erupt + scale_fill_gradientn(colours = colorspace::diverge_hcl(7))

erupt + scale_fill_distiller()
erupt + scale_fill_distiller(palette = "RdPu")
erupt + scale_fill_distiller(palette = "YlOrBr")
```  

```{r}
df <- data.frame(x = 1, y = 1:5, z = c(1, 3, 2, NA, 5))
p <- ggplot(df, aes(x, y)) + geom_tile(aes(fill = z), size = 5)
p
# Make missing colours invisible
p + scale_fill_gradient(na.value = NA)
# Customise on a black and white scale
p + scale_fill_gradient(low = "black", high = "white", na.value = "red")
```  

##### 6.6.2.2 Discrete  
```{r}
df <- data.frame(x = c("a", "b", "c", "d"), y = c(3, 4, 1, 2))
bars <- ggplot(df, aes(x, y, fill = x)) + 
  geom_bar(stat = "identity") +
  labs(x = NULL, y = NULL) + 
  theme(legend.position = "none")

bars
bars + scale_fill_hue(c = 40)
bars + scale_fill_hue(h = c(180, 300))

bars + scale_fill_brewer(palette = "Set1")
bars + scale_fill_brewer(palette = "Set2")
bars + scale_fill_brewer(palette = "Accent")

bars + scale_fill_grey()
bars + scale_fill_grey(start = 0.5, end = 1)
bars + scale_fill_grey(start = 0, end = 0.5)


```  


```{r}
#Bright colours work best with points
df <- data.frame(x = 1:3 + runif(30), y = runif(30), z = c("a", "b", "c"))
point <- ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z)) +
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)
point + scale_colour_brewer(palette = "Set1")
point + scale_colour_brewer(palette = "Set2")
point + scale_colour_brewer(palette = "Pastel1")


# Subtler colours work better with areas
df <- data.frame(x = 1:3, y = 3:1, z = c("a", "b", "c"))
area <- ggplot(df, aes(x, y)) +
  geom_bar(aes(fill = z), stat = "identity") +
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)
area + scale_fill_brewer(palette = "Set1")
area + scale_fill_brewer(palette = "Set2")
area + scale_fill_brewer(palette = "Pastel1")
```  


#### 6.6.3 The manual discrete scale  
```{r}
plot <- ggplot(msleep, aes(brainwt, bodywt)) +
  scale_x_log10() +
  scale_y_log10()
plot
plot +
  geom_point(aes(colour = vore)) +
  scale_colour_manual(
    values = c("red", "orange", "green", "blue"),
    na.value = "grey50"
    )

colours <- c(
  carni = "red",
  insecti = "orange",
  herbi = "green",
  omni = "blue"
  )
plot +
  geom_point(aes(colour = vore)) +
  scale_colour_manual(values = colours)


huron <- data.frame(year = 1875:1972, level = as.numeric(LakeHuron))
ggplot(huron, aes(year)) +
  geom_line(aes(y = level + 5), colour = "red") +
  geom_line(aes(y = level - 5), colour = "blue")


ggplot(huron, aes(year)) +
  geom_line(aes(y = level + 5, colour = "above")) +
  geom_line(aes(y = level - 5, colour = "below"))
  

ggplot(huron, aes(year)) +
  geom_line(aes(y = level + 5, colour = "above")) +
  geom_line(aes(y = level - 5, colour = "below")) +
  scale_colour_manual("Direction",
                      values = c("above" = "red", "below" =
                                   "blue")
                      )
```



#### 6.6.4 The identity scale  

```{r}
ggplot(luv_colours, aes(u, v)) +
  geom_point(aes(colour = col), size = 3) +
  scale_color_identity() +
  coord_equal()
```  


## Chapter 7 Positioning  
### Introduction  

### 7.2 Faceting  
`facet_null()` a single plot  

#### 7.2.1 `facet_wrap()`  
`as.table` controls whether the facets are laid out like a table (TRUE), with highest values at the bottom-right, or a plot (FALSE), with the highest values at the
top-right. `dir` controls the direction of wrap: horizontal or vertical.

```{r}
library(ggplot2)
mpg2 <- subset(mpg, cyl != 5 & drv %in% c("4", "f") & class != "2seater")
base <- ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL)
base + facet_wrap(~class, ncol = 3)
base + facet_wrap(~class, ncol = 3, as.table = FALSE)

base + facet_wrap(~class, nrow = 3)
base + facet_wrap(~class, nrow = 3, dir = "v")
```  

#### 7.2.2 `facet_grid()`  
```{r}
base + facet_grid(. ~ cyl)
base + facet_grid(drv ~ .)
base + facet_grid(drv ~ cyl)
```  

#### 7.2.3 Controlling scales  

```{r}
p <- ggplot(mpg2, aes(cty, hwy)) +
  geom_abline() +
  geom_jitter(width = 0.1, height = 0.1)
p + facet_wrap(~cyl)
p + facet_wrap(~cyl, scales = "free")
```  

```{r}
economics_long
ggplot(economics_long, aes(date, value)) + 
  geom_line() +
  facet_wrap(~variable, scales = "free_y", ncol = 1)


mpg2$model <- reorder(mpg2$model, mpg2$cty)
mpg2$manufacturer <- reorder(mpg2$manufacturer,-mpg2$cty)
ggplot(mpg2, aes(cty, model)) +
  geom_point() +
  facet_grid(manufacturer ~ ., scales = "free", space = "free") +
  theme(strip.text.y = element_text(angle = 0))
```  


#### 7.2.4 Missing faceting variables  

```{r}
df1 <- data.frame(x = 1:3, y = 1:3, gender = c("f", "f", "m"))
df2 <- data.frame(x = 2, y = 2)
ggplot(df1, aes(x, y)) +
  geom_point(data = df2, colour = "red", size = 2) +
  geom_point() +
  facet_wrap(~gender)
```  


#### 7.2.5 Gruoping vs. faceting  
```{r}
library(dplyr)
df <- data.frame(
  x = rnorm(120, c(0, 2, 4)),
  y = rnorm(120, c(1, 2, 1)),
  z = letters[1:3]
)
ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z))

ggplot(df, aes(x, y)) +
  geom_point() +
  facet_wrap(~z)

df_sum <- df %>%
  group_by(z) %>%
  summarise(x = mean(x), y = mean(y)) %>%
  rename(z2 = z)

ggplot(df, aes(x, y)) +
  geom_point() +
  geom_point(aes(colour = z2), data = df_sum, size = 4) +
  facet_wrap(~z)
# put all the data in the background of each panel:
df2 <- select(df, -z)
ggplot(df, aes(x, y)) +
  geom_point(data = df2, colour = "grey70") +
  geom_point(aes(colour = z)) +
  facet_wrap(~z)
```  

#### 7.2.6 Continuous variables  

```{r}
# Bins of width 1
mpg2$disp_w <- cut_width(mpg2$displ, 1)
# Six bins of equal length
mpg2$disp_i <- cut_interval(mpg2$displ, 6)
# Six bins containing equal numbers of points
mpg2$disp_n <- cut_number(mpg2$displ, 6)

plot <- ggplot(mpg2, aes(cty, hwy)) +
  geom_point() +
  labs(x = NULL, y = NULL)
plot + facet_wrap(~disp_w, nrow = 1)
plot + facet_wrap(~disp_i, nrow = 1)
plot + facet_wrap(~disp_n, nrow = 1)
```   


### 7.3 Coordinate systems  
There are two types of coordinate system. Linear coordinate system preserve the shape of geoms:  

`coord_cartesian()`: the default Cartesian coordinate system, where the 2d position of an element is given by the combination of the x and y positions.  

`coord_fixed()`: Cartesian coordinate system with a fixed aspect ratio.  

`coord_flip()`: Cartesian coordinate system with x and y axes flipped.  

On the other hand, non-linear coordinate can change the shapes: a straight line may no longer be straight. The closest distance between two points is no longer be a straight line.  

`coord_map()/coord_quickmap()`: Map projections.  

`coord_polar()`: Polar coordinates.  

`coord_trans()`: Apply arbitrary transformations to x and y positions, after the data has been processed by the stat.  

### 7.4 Linear coordinate systems  
#### 7.4.1 Zooming into a plot with `coord_cartesian()`  

The key difference is how the limits work: when setting scale limits, any data outside the limits is thrown away; but when setting coordinate system limits we still use all the data, but we only display a small region of the plot. Setting coordinate system limits is like looking at the plot under a magnifying glass.  

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
# Full dataset
base
# Scaling to 5--7 throws away data outside that range
base + scale_x_continuous(limits = c(5, 7))
# Zooming to 5--7 keeps all the data but only shows some of it
base + coord_cartesian(xlim = c(5, 7))
```  

#### 7.4.2 Flipping the axes with `coord_flip()`  
```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  geom_smooth()
# Exchanging cty and displ rotates the plot 90 degrees, but the smooth
# is fit to the rotated data.
ggplot(mpg, aes(cty, displ)) +
  geom_point() +
  geom_smooth()

# coord_flip() fits the smooth to the original data, and then rotates
# the output
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  geom_smooth() +
  coord_flip()

```  

#### 7.4.3 Equal scales with `coord_equal()`  

`coord_equal()` fixes the ratio of length on the x and y axes. The default ratio, ensures that the x and y axes have equal scales  

### 7.5 Non-linear coordinate systems  
```{r}
rect <- data.frame(x = 50, y = 50)
line <- data.frame(x = c(1, 200), y = c(100, 1))
base <- ggplot(mapping = aes(x, y)) +
  geom_tile(data = rect, aes(width = 50, height = 50)) +
  geom_line(data = line) +
  xlab(NULL) + ylab(NULL)
base
base + coord_polar("x")
base + coord_polar("y")
base + coord_flip()
base + coord_trans(y = "log10")
base + coord_equal()
```  

Firstly, the parameterisation of each geom is changed to
be purely location-based, rather than location and dimension based.  


```{r}
df <- data.frame(r = c(0, 1), theta = c(0, 3 / 2 * pi))
ggplot(df, aes(r, theta)) +
  geom_line() +
  geom_point(size = 2, colour = "red")

interp <- function(rng, n) {
  seq(rng[1], rng[2], length = n)
  }

munched <- data.frame(
  r = interp(df$r, 15),
  theta = interp(df$theta, 15)
  )
ggplot(munched, aes(r, theta)) +
  geom_line() +
  geom_point(size = 2, colour = "red")

transformed <- transform(munched,
                         x = r * sin(theta),
                         y = r * cos(theta)
                         )

 ggplot(transformed, aes(x, y)) +
     geom_path() +
     geom_point(size = 2, colour = "red") +
     coord_equal()

```  


#### 7.5.1 Transformations with `coord_trans()`  
```{r}
# Linear model on original scale is poor fit
base <- ggplot(diamonds, aes(carat, price)) +
  stat_bin2d() +
  geom_smooth(method = "lm") +
  xlab(NULL) +
  ylab(NULL) +
  theme(legend.position = "none")
base
# Better fit on log scale, but harder to interpret
base +
  scale_x_log10() +
  scale_y_log10()
# Fit on log scale, then backtransform to original.
# Highlights lack of expensive diamonds with large carats
pow10 <- scales::exp_trans(10)
base +
  scale_x_log10() +
  scale_y_log10() +
  coord_trans(x = pow10, y = pow10)
```   


#### 7.5.2 Polor coordinates with `coord_polar()`  

```{r}
base <- ggplot(mtcars, aes(factor(1), fill = factor(cyl))) +
  geom_bar(width = 1) +
  theme(legend.position = "none") +
  scale_x_discrete(NULL, expand = c(0, 0)) +
  scale_y_continuous(NULL, expand = c(0, 0))
# Stacked barchart
base

# Pie chart
base + coord_polar(theta = "y")
# The bullseye chart
base + coord_polar()
```  


#### 7.5.3 Map projections with `coord_map()`  
```{r}
#library(map_data)
# Prepare a map of NZ
nzmap <- ggplot(map_data("nz"), aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  xlab(NULL) + 
  ylab(NULL)
# Plot it in cartesian coordinates
nzmap
# With the aspect ratio approximation
nzmap + coord_quickmap()


world <- map_data("world")
worldmap <- ggplot(world, aes(long, lat, group = group)) +
  geom_path() +
  scale_y_continuous(NULL, breaks = (-2:3) * 30, labels = NULL) +
  scale_x_continuous(NULL, breaks = (-4:4) * 45, labels = NULL)


worldmap + coord_map()
   # Some crazier projections
worldmap + coord_map("ortho")
worldmap + coord_map("stereographic")
```




