---
title: "ggplot2 Chapter1-3"
author: "Li Sun"
date: "September 6, 2017"
output: html_document
---
## Chapter1  
### 1.3 What is the grammar of graphics?  
The grammar tells us that a statistical graphic is a mapping from data to aesthetic attributes of geometric objects.  
* Data  
* Layers, made up of geomtric elements and statistical transformation.  
* Scales  
* Coordinate system  
* Facet: how to break up the data into subsets and display those subsets as small multiples.  
* Theme: controls the finer points of display, font size and background colour.  
* ggplot2 can only create static graphics.  


## Chapter 2 Getting started with ggplot2  
###2.1 Introduction  
###2.2 Fuel economy data  
```{r}
library(ggplot2)
library(dplyr)
#library(magrittr)
mpg
```  

* `cty` and hwy record miles per gallon (mpg) for city and highway driving.  

* `displ` is the engine displacement in litres.  

* `drv` is the drive train: front wheel (f), rear wheel (r) or four wheel (4).  

* `model` is the model of car. There are 38 models, selected because they had a new edition every year between 1999 and 2008.  

* `class` (not shown), is a categorical variable describing the “type” of car: two seater, SUV, compact, etc.  


### 2.3 Key components  
1. data
2. Aesthetic mappings between variables in data and visual properties.  
3. Layer(s)  

```{r}
ggplot(mpg,aes(x=displ, y=hwy))+
  geom_point()
```  
data: mpg  
aesthetic mapping: engine size mapped to x position, fuel economy to y position.  
layer: points  
__data__ and __aesthetic__ mappings are supplied in ggplot(), then layers are added on with +.   

### 2.4 Color, size, shape and other aesthetic attributes.  
```{r}
ggplot(mpg, aes(displ, cty, colour = class)) +
  geom_point()
```  
This gives each point a unique colour corresponding to its class.  
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue"))
ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue")
```  

In the first plot, the value “blue” is scaled to a pinkish colour, and a legend is added. In the second plot, the points are given the R colour blue.  
When using aesthetics in a plot, less is usually more.  

### 2.5 Faceting  
Faceting creates tables of graphics by splitting the data into subsets and displaying the same graph for each subset.  
Wrapped is the most useful, so we’ll discuss it here, grid facetting later.  
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class)
```  

### 2.6 Plot geoms  
#### 2.6.1 Adding a smoother to a plot  
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth() #se = TRUE
```  

This overlays the scatterplot with a smooth curve, including an assessment of uncertainty in the form of point-wise confidence interval shown in grey.  
* `method = 'loess'`, default for small n, uses a smooth local regression, the wiggliness of the line is controlled by the `span` parameter (0,1).  
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 0.2)
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 1)
```  

* `method = 'gam'` fits a generalised additive model provided by the mgcv package.  
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "gam", formula = y ~ s(x))
```  

* `method = 'lm' fits a linear mdoel.  
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```  

* `method = "rlm"` works like lm(), but uses a robust fitting algorithm so that outliers don’t affect the fit as much.
```{r}
library(MASS)
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "rlm")
```  

#### 2.6.2 Boxplots and jittered points  
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_point()
```  

Jittering, `geom_jitter()`, adds a little random noise to the data which can help avoid overplotting.  
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_jitter()
```  
Boxplots, `geom_boxplot()`, summarise the shape of the distribution with a handful of summary statistics.  
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_boxplot()
```  
Violin plots, `geom_violin()`, show a compact representation of the “density” of the distribution, highlighting the areas where more points are found.  
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_violin()
```  

#### 2.6.3 Histograms and frequency polygons  
```{r}
ggplot(mpg, aes(hwy)) + geom_histogram()
```  
```{r}
ggplot(mpg, aes(hwy)) + geom_freqpoly()
```  

You should always try many bin widths, and you may find you need multiple bin widths to tell the full story of your data.  
```{r}
ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 2.5)
```
```{r}
ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 1)
```  

I’m not a fan of density plots because they are harder
to interpret since the underlying computations are more complex.  

```{r}
ggplot(mpg, aes(hwy)) + 
  geom_density(binwidth = 1)
```  

```{r}
ggplot(mpg, aes(displ, colour = drv)) + 
  geom_freqpoly(binwidth = 0.5)
ggplot(mpg, aes(displ, fill = drv)) + 
  geom_histogram(binwidth = 0.5) + 
  facet_wrap(~drv, ncol = 1)
```  

#### 2.6.4 Bar charts  
unsummarised data:  

```{r}
ggplot(mpg, aes(manufacturer)) + 
  geom_bar() 
```  

Presumarised data:  
```{r}
drugs <- data.frame(
  drug = c("a", "b", "c"),
  effect = c(4.2, 9.7, 6.1)
  )

ggplot(drugs, aes(drug, effect)) + 
  geom_bar(stat = "identity")
ggplot(drugs, aes(drug, effect)) + geom_point()
```  

#### 2.6.5 Time series with line and path plots  
Line plots usually have time on the x-axis, showing how
a single variable has changed over time. 
```{r}
ggplot(economics, aes(date, unemploy / pop)) + 
  geom_line()

ggplot(economics, aes(date, uempmed)) + 
  geom_line()
```  

Path plots show how two variables have simultaneously changed over time, with time encoded in the way that observations are connected.  
```{r}
ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path() + 
  geom_point()
```  
Because of the many line crossings, the direction in which time flows isn’t easy to see in the first plot.  

```{r}
year <- function(x) as.POSIXlt(x)$year + 1900
ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path(colour = "grey50") + 
  geom_point(aes(colour = year(date)))
```  

### 2.7 Modifying the axes  
```{r}
ggplot(mpg, aes(cty, hwy)) + 
  geom_point(alpha = 1 / 3)

ggplot(mpg, aes(cty, hwy)) + 
  geom_point(alpha = 1 / 3) + 
  xlab("city driving (mpg)") + 
  ylab("highway driving (mpg)")

# Remove the axis labels with NULL
ggplot(mpg, aes(cty, hwy)) + 
  geom_point(alpha = 1 / 3) + 
  xlab(NULL) + 
  ylab(NULL)

```   

xlim() and ylim() modify the limits of axes:  
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_jitter(width = 0.25) 
ggplot(mpg, aes(drv, hwy)) + 
  geom_jitter(width = 0.25) + 
  xlim("f", "r") + 
  ylim(20, 30)
ggplot(mpg, aes(drv, hwy)) + 
  geom_jitter(width = 0.25, na.rm = TRUE) + 
  ylim(NA, 30)
```  

You can suppress the associated warning with na.rm = TRUE.  

### 2.8 Output  
```{r}
p <- ggplot(mpg, aes(displ, hwy, colour = factor(cyl))) + 
  geom_point()
```  
Render it on screen, with print().  
```{r}
print(p)
```  
Save it to disk, with ggsave()  
```{r}
ggsave("plot.png", width = 5, height = 5)
```  
Briefly describe its structure with summary().  
```{r}
summary(p)
```  
Save a cached copy of it to disk, with `saveRDS()`. This saves a complete copy of the plot object, so you can easily re-create it with `readRDS()`.  

### 2.9 Quick plots  
```{r}
qplot(displ, hwy, data = mpg)
qplot(displ, data = mpg)
```  
Unless otherwise specified, qplot() tries to pick a sensible geometry and statistic based on the arguments provided.  

If you want to set an aesthetic to a constant, you need to use `I()`:  
```{r}
qplot(displ, hwy, data = mpg, colour = "blue")
qplot(displ, hwy, data = mpg, colour = I("blue"))
```




## Chapter3 Toolbox  
### 3.1 Introduction  
### 3.2 Basic plot types  
Each of these geoms is two dimensional and requires both x and y aesthetics.  
All of them understand colour (or color) and size aesthetics, and the filled geoms (bar, tile and polygon) also understand fill.  
```{r}
df <- data.frame(
  x = c(3, 1, 5),
  y = c(2, 4, 6),
  label = c("a","b","c")
  )
p <- ggplot(df, aes(x, y, label = label)) + 
  labs(x = NULL, y = NULL) + # Hide axis label 
  theme(plot.title = element_text(size = 12)) # Shrink plot title
```   

* geom_area() draws an area plot, which is a line plot filled to the y-axis (filled lines). Multiple groups will be stacked on top of each other.  
```{r}
p + geom_area() + ggtitle("area")
```  

* `geom_bar(stat = "identity")`makes a barchart. We need `stat = "identity"` because the default stat automatically counts values (so is essentially a 1d geom. The identity stat leaves the data unchanged. Multiple bars in the same location will be stacked on top of one another.  
```{r}
p + geom_bar(stat = "identity") + ggtitle("bar")
```  

* `geom_line()` makes a line plot. The group aesthetic determines which observations are connected  
```{r}
p + geom_line() + ggtitle("line")
```  

* geom_path() is similar to a geom_line(), but lines are connected in the order they appear in the data, not from left to right.  
```{r}
p + geom_path() + ggtitle("path")
```  

* `geom_point()` produces a scatterplot. geom_point() also understands the shape aesthetic.  

```{r}
p + geom_point() + ggtitle("point")
```  

* geom_polygon() draws polygons, which are filled paths.  

```{r}
p + geom_polygon() + ggtitle("polygon")
```  

* `geom_rect()` is parameterised by the four corners of the rectangle, `xmin, ymin, xmax and ymax`. `geom_tile()` is exactly the same, but parameterised by the center of the rect and its size, `x, y, width and height`.  
`geom_raster()` is a fast special case of `geom_tile()` used when all the tiles are the same size.  

```{r}
p + geom_tile() + ggtitle("raster")
```  

### 3.3 Lables  
`geom_text()` has the most aesthetics of any geom  
* `family` gives the name of a font.  
```{r}
df <- data.frame(x = 1, y = 3:1, family = c("sans", "serif", "mono"))
ggplot(df, aes(x, y)) + 
  geom_text(aes(label = family, family = family))
```  

* `fontface` specifies the face: “plain” (the default), “bold” or “italic”.  
```{r}
df <- data.frame(x = 1, y = 3:1, face = c("plain", "bold", "italic"))
ggplot(df, aes(x, y)) + 
  geom_text(aes(label = face, fontface = face))
```  

You can adjust the alignment of the text with the hjust (“left”, “center”, “right”, “inward”, “outward”) and vjust (“bottom”, “middle”, “top”, “inward”, “outward”) aesthetics.  
```{r}
df <- data.frame( 
  x = c(1, 1, 2, 2, 1.5), 
  y = c(1, 2, 1, 2, 1.5), 
  text = c( 
    "bottom-left", "bottom-right", 
    "top-left", "top-right", "center"
    )
  )
ggplot(df, aes(x, y)) + 
  geom_text(aes(label = text)) 
ggplot(df, aes(x, y)) + 
  geom_text(aes(label = text), vjust = "inward", hjust = 
              "inward")
```  


* `size` controls the font size.  
* `angle` specifies the rotation of the text in degrees.  

* The `nudge_x` and `nudge_y` parameters allow you to nudge the text a little horizontally or vertically:  
```{r}
df <- data.frame(trt = c("a", "b", "c"), resp = c(1.2, 3.4, 2.5))
ggplot(df, aes(resp, trt)) + 
  geom_point() + 
  geom_text(aes(label = paste0("(", resp, ")")), nudge_y = -0.25) + 
  xlim(1, 3.6)
```  

* If `check_overlap = TRUE`, overlapping labels will be automatically removed.  

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_text(aes(label = model)) + 
  xlim(1, 8)
ggplot(mpg, aes(displ, hwy)) +
  geom_text(aes(label = model), check_overlap = TRUE) +
  xlim(1, 8)
```  


A variation on `geom_text()` is `geom_label()`: it draws a rounded rectangle behind the text. This makes it useful for adding labels to plots with busy backgrounds:  
```{r}
label <- data.frame(
  waiting = c(55, 80),
  eruptions = c(2, 4.3),
  label = c("peak one", "peak two")
  )
ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_tile(aes(fill = density)) +
  geom_label(data = label, aes(label = label))
```  

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point()
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = class), method =
                          "smart.grid")
```  


### 3.4 Annotations  
* `geom_text()` to add text descriptions or to label points  

* `geom_rect()` to highlight interesting rectangular regions of the plot.  

* `geom_line()`, `geom_path()` and `geom_segment()` to add lines.  
```{r}
ggplot(economics, aes(date, unemploy)) + 
  geom_line()
```  

* `geom_vline()`, `geom_hline()` and `geom_abline()` allow you to add reference lines (sometimes called rules), that span the full range of the plot.  
```{r}
presidential <- subset(presidential, start > economics$date[1])
ggplot(economics) + 
  geom_rect(
    aes(xmin = start, xmax = end, fill = party),
    ymin = -Inf, ymax = Inf, alpha = 0.2,
    data = presidential
    ) +
  geom_vline(
    aes(xintercept = as.numeric(start)),
    data = presidential,
    colour = "grey50", alpha = 0.5
    ) +
  geom_text(
    aes(x = start, y = 2500, label = name),
    data = presidential,
    size = 3, vjust = 0, hjust = 0, nudge_x = 50
    ) +
  geom_line(aes(date, unemploy)) +
  scale_fill_manual(values = c("blue", "red"))
```  

```{r}
yrng <- range(economics$unemploy)
xrng <- range(economics$date)
caption <- paste(strwrap("Unemployment rates in the US have
                         varied a lot over the years", 40),
                 collapse = "\n")
ggplot(economics, aes(date, unemploy)) +
  geom_line() +
  geom_text(
    aes(x, y, label = caption),
    data = data.frame(x = xrng[1], y = yrng[2], caption = caption),
    hjust = 0, vjust = 1, size = 4
    )
```  

It’s easier to use the `annotate()` helper function which creates the data frame for you:  
```{r}
ggplot(economics, aes(date, unemploy)) + 
  geom_line() + 
  annotate("text", x = xrng[1], y = yrng[2], label = caption,hjust = 0, vjust = 1, size = 4 
           )
```  

it’s much easier to see the subtle differences if we add a reference line.  

```{r}
ggplot(diamonds, aes(log10(carat), log10(price))) + 
  geom_bin2d() + 
  facet_wrap(~cut, nrow = 1)


mod_coef <- coef(lm(log10(price) ~ log10(carat), data = diamonds))
ggplot(diamonds, aes(log10(carat), log10(price))) + 
  geom_bin2d() + 
  geom_abline(intercept = mod_coef[1], slope = mod_coef[2], 
              colour = "white", size = 1) + 
  facet_wrap(~cut, nrow = 1)
```  

### 3.5 Collective geoms  
An __individual__ geom draws a distinct graphical object for each observation (row). For example, the point geom draws one point per row. A __collective__ geom displays multiple observations with one geometric object.  

By default, the `group` aesthetic is mapped to the interaction of all discrete variables in the plot.  
```{r}
data(Oxboys, package = "nlme")
head(Oxboys)
```  

#### 3.5.1 Multiple groups, one aesthetic  
you want to be able to distinguish individual subjects, but not identify them.  
```{r}
ggplot(Oxboys, aes(age, height, group = Subject)) +
  geom_point() + 
  geom_line()
```   

Incorrect:  
```{r}
ggplot(Oxboys, aes(age, height)) + 
  geom_point() + 
  geom_line()
```  

#### 3.5.2 Different groups on different layers  

showing the overall trend for all boys.  
```{r}
ggplot(Oxboys, aes(age, height, group = Subject)) + 
  geom_line() + 
  geom_smooth(method = "lm", se = FALSE)
```  

Instead of setting the grouping aesthetic in ggplot(), where it will apply to all layers, we set it in geom_line() so it applies only to the lines.  

```{r}
ggplot(Oxboys, aes(age, height)) + 
  geom_line(aes(group = Subject)) + 
  geom_smooth(method = "lm", size = 2, se = FALSE)
```  

#### 3.5.3 Overriding the default grouping  
```{r}
ggplot(Oxboys, aes(Occasion, height)) + 
  geom_boxplot()
```  

Now we want to overlay lines that connect each individual boy. Simply adding geom_line() does not work: the lines are drawn within each occassion, not across each subject.  
```{r}
ggplot(Oxboys, aes(Occasion, height)) + 
  geom_boxplot() + 
  geom_line(colour = "#3366FF", alpha = 0.5)
```

we need to override the grouping to say we want one line per boy:  
```{r}
ggplot(Oxboys, aes(Occasion, height)) + 
  geom_boxplot() + 
  geom_line(aes(group = Subject), colour = "#3366FF", alpha = 0.5)
```  

#### 3.5.4 Matching aesthetics to graphic objects  
There is one more observation than line segment, and so the aesthetic for the first observation is used for the first segment, the second observation for the second segment
and so on.This means that the aesthetic for the last observation is not used:  
```{r}
df <- data.frame(x = 1:3, y = 1:3, colour = c(1,3,5))
ggplot(df, aes(x, y, colour = factor(colour))) +
  geom_line(aes(group = 1), size = 2) + 
  geom_point(size = 5)
ggplot(df, aes(x, y, colour = colour)) +
  geom_line(aes(group = 1), size = 2) +
  geom_point(size = 5)
```  
you can perform the linear interpolation yourself:  
```{r}
xgrid <- with(df, seq(min(x), max(x), length = 50))
interp <- data.frame( 
  x = xgrid,
  y = approx(df$x, df$y, xout = xgrid)$y,
  colour = approx(df$x, df$colour, xout = xgrid)$y
  )
ggplot(interp, aes(x, y, colour = colour)) +
  geom_line(size = 2) +
  geom_point(data = df, size = 5)
```   


how would you colour a polygon that had a different fill colour for each point on its border?  
```{r}
ggplot(mpg, aes(class)) + 
  geom_bar()
ggplot(mpg, aes(class, fill = drv)) + 
  geom_bar()
```  

If you try to map fill to a continuous variable in the same way, it doesn’t work.  
To show multiple colours, we need multiple bars for each class, which we can get by overriding the grouping:  
```{r}
ggplot(mpg, aes(class, fill = hwy)) + 
  geom_bar()
ggplot(mpg, aes(class, fill = hwy, group = hwy)) +
  geom_bar()
```  

### 3.6 Surface plots  
```{r}
ggplot(faithfuld, aes(eruptions, waiting)) + 
  geom_contour(aes(z = density, colour = ..level..))
ggplot(faithfuld, aes(eruptions, waiting))+
  geom_raster(aes(fill = density))
```    

Bubble plots work better with fewer observations
```{r}
small <- faithfuld[seq(1, nrow(faithfuld), by = 10), ]
ggplot(small, aes(eruptions, waiting)) +
  geom_point(aes(size = density), alpha = 1/3) +
  scale_size_area()
```  


### 3.7 Drawing maps  
#### 3.7.1 Vector boundaries  
Vector boundaries are defined by a data frame with one row for each “corner” of a geographical region like a country, state, or county. It requires four variables:  
* `lat` and `long`, giving the location of a point.  
* `group`, a unique identifier for each contiguous region.  
* `id`, the name of the region.  

```{r}
#mi_counties <- ggplot2::map_data("county", "michigan") %>% 
#  select(lon = long, lat, group, id = subregion)
#head(mi_counties)
#ggplot(mi_counties, aes(lon, lat)) +
#  geom_polygon(aes(group = group)) +
#  coord_quickmap()
#ggplot(mi_counties, aes(lon, lat)) +
#  geom_polygon(aes(group = group), fill = NA, colour = "grey50") +
#  coord_quickmap()
```  


#### 3.7.2 Point metadata  
```{r}
#mi_cities <- maps::us.cities %>%
#  tbl_df() %>%
#  filter(country.etc == "MI") %>%
#  select(-country.etc, lon = long) %>%
#  arrange(desc(pop))
#mi_cities
```  
It’s not terribly useful without a reference. You almost always combine point metadata with another layer to make it interpretable.  

```{r}
#ggplot(mi_cities, aes(lon, lat)) +
#  geom_point(aes(size = pop)) +
#  scale_size_area() +
#  coord_quickmap()
#ggplot(mi_cities, aes(lon, lat)) +
#  geom_polygon(aes(group = group), mi_counties, fill = NA, colour = "grey50") +
#  geom_point(aes(size = pop), colour = "red") +
#  scale_size_area() +
#  coord_quickmap()
```  

#### 3.7.3 Raster images  

### 3.8 Revealing uncertainty  
* Discrete x, range: `geom_errorbar`, `geom_linerange()` 
```{r}
y <- c(18, 11, 16)
df <- data.frame(x = 1:3, y = y, se = c(1.2, 0.5, 1.0))
base <- ggplot(df, aes(x, y, ymin = y - se, ymax = y + se))
base + geom_errorbar()
base + geom_linerange()
```
* Discrete x, range & center:  `geom_crossbar()`,`geom_pointrange()`  
```{r}
base + geom_crossbar()
base + geom_pointrange()
```  

* Continuous x, range: `geom_ribbon()`  
```{r}
base + geom_ribbon()
```  

* Continuous x, range & center: `geom_smooth(stat = "identity")`  
```{r}
base + geom_smooth(stat = "identity")
```  


### 3.9 Weighted data  
There are two aesthetic attributes that can be used to adjust for weights. Firstly, for simple geoms like lines and points, use the size aesthetic:  

Unweighted  
```{r}
ggplot(midwest, aes(percwhite, percbelowpoverty)) + 
  geom_point()
```  
Weight by population  
```{r}
ggplot(midwest, aes(percwhite, percbelowpoverty)) +
  geom_point(aes(size = poptotal / 1e6)) +
  scale_size_area("Population\n(millions)", breaks = c(0.5, 1, 2, 4))
```

These weights will be passed on to the statistical summary function.  
Unweighted
```{r}
ggplot(midwest, aes(percwhite, percbelowpoverty)) +
  geom_point() +
  geom_smooth(method = lm, size = 1)
```  

Weighted by population  
```{r}
ggplot(midwest, aes(percwhite, percbelowpoverty)) +
  geom_point(aes(size = poptotal / 1e6)) +
  geom_smooth(aes(weight = poptotal), method = lm, size = 1) +
  scale_size_area(guide = "none")
```  

The following code shows the difference this makes for a histogram of the percentage below the poverty line:  
```{r}
ggplot(midwest, aes(percbelowpoverty)) +
  geom_histogram(binwidth = 1) +
  ylab("Counties")
ggplot(midwest, aes(percbelowpoverty)) +
  geom_histogram(aes(weight = poptotal), binwidth = 1) +
  ylab("Population (1000s)")
```  

### 3.10 Diamonds data  
```{r}
diamonds
```  

### 3.11 Displaying distributions  
For 1d continuous distributions the most important geom is the histogram, `geom_histogram()`:  
```{r}
ggplot(diamonds, aes(depth)) +
  geom_histogram()

ggplot(diamonds, aes(depth)) +
  geom_histogram(binwidth = 0.1) +
  xlim(55, 70)
```  

If you want to compare the distribution between groups, you have a few options:  
* Show small multiples of the histogram, facet_wrap(~ var).  

* Use colour and a frequency polygon, geom_freqpoly().  
TODO: add na.rm back  
```{r}
ggplot(diamonds, aes(depth)) +
  geom_freqpoly(aes(colour = cut), binwidth = 0.1) +
  xlim(58, 68) +
  theme(legend.position = "none")
```

* Use a “conditional density plot”, geom_histogram(position = "fill").  
```{r}
ggplot(diamonds, aes(depth)) + 
  geom_histogram(aes(fill = cut), binwidth = 0.1, position = "fill") +
  xlim(58, 68) +
  theme()
```  

An alternative to a bin-based visualisation is a density estimate.  
`geom_density(`) places a little normal distribution at each data point and sums up all the curves.  
```{r}
ggplot(diamonds, aes(depth)) + 
  geom_density(na.rm = TRUE) +
  xlim(58, 68) +
  theme()

ggplot(diamonds, aes(depth, fill = cut, colour = cut)) +
  geom_density(alpha = 0.2, na.rm = TRUE) +
  xlim(58, 68) +
  theme()
```  

sometimes you want to compare many distributions, and it’s useful to have alternative options that sacrifice quality for
quantity. Here are three options:
* `geom_boxplot()`: the box-and-whisker plot shows five summary statistics along with individual “outliers”.  
```{r}
ggplot(diamonds, aes(clarity, depth)) + 
  geom_boxplot()
ggplot(diamonds, aes(carat, depth)) +
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  xlim(NA, 2.05)
```  

* `geom_violin()`: the violin plot is a compact version of the density plot.  
```{r}
ggplot(diamonds, aes(clarity, depth)) + 
  geom_violin()
ggplot(diamonds, aes(carat, depth)) +
  geom_violin(aes(group = cut_width(carat, 0.1))) +
  xlim(NA, 2.05)
```  

* `geom_dotplot()`: draws one point for each observation  


### 3.12 Dealing with overplotting  
* Very small amounts of overplotting can sometimes be alleviated by making the points smaller, or using hollow glyphs.  
```{r}
df <- data.frame(x = rnorm(2000), y = rnorm(2000))
norm <- ggplot(df, aes(x, y)) + xlab(NULL) + ylab(NULL)
norm + geom_point()
norm + geom_point(shape = 1) # Hollow circles
norm + geom_point(shape = ".") # Pixel sized
```  

* For larger datasets with more overplotting, you can use alpha blending (transparency) to make the points transparent.
  
  
```{r}
norm + geom_point(alpha = 1 / 3)
norm + geom_point(alpha = 1 / 5)
norm + geom_point(alpha = 1 / 10)
```  

* you can randomly jitter the points to alleviate some overlaps with `geom_jitter()`.  

* Bin the points and count the number in each bin, then visualise that count (the 2d generalisation of the histogram), geom_bin2d().  
```{r}
norm + geom_bin2d()
norm + geom_bin2d(bins = 10)
norm + geom_hex()
norm + geom_hex(bins = 10)
```   

### 3.13 Statistical summaries  

how we can count the number of diamonds in each bin:  
```{r}
ggplot(diamonds, aes(color)) + 
  geom_bar()
ggplot(diamonds, aes(color, price)) +
  geom_bar(stat = "summary_bin", fun.y = mean)
```  

add na.rm back  
```{r}
ggplot(diamonds, aes(table, depth)) + 
  geom_bin2d(binwidth = 1) +
  xlim(50, 70) +
  ylim(50, 70)
ggplot(diamonds, aes(table, depth, z = price)) +
  geom_raster(binwidth = 1, stat = "summary_2d", fun = mean) +
  xlim(50, 70) +
  ylim(50, 70)

```



































